Dysregulation of Cerebral Hemoglobin Content in Malawian Children with Malaria

### 2021-05-13 figures.Rmd
table 1. changed range to quartiles figure 1. split alpha into fig 2

# Age, hematocrit, and lactate
df_clinical_res <- df_hbox %>%

@@

  group_by(Status) %>% 
  summarise(`Age (years)` = median(Age, na.rm = TRUE), age_range = range(Age, na.rm = TRUE),

            `Temperature (°C)` = median(Temperature, na.rm = TRUE), temp_range = range(Temperature, na.rm = TRUE),
            `Heart rate` = median(HR, na.rm = TRUE), hr_range = range(HR, na.rm = TRUE),
            `Systolic BP` = median(BP.Systolic, na.rm = TRUE), sys_range = range(BP.Systolic, na.rm = TRUE),
            `Diastolic BP` = median(BP.Diastolic, na.rm = TRUE), dia_range = range(BP.Diastolic, na.rm = TRUE),
            `Respiration rate` = median(RR, na.rm = TRUE), rr_range = range(RR, na.rm = TRUE),
            `O~2~ saturation` = median(O2.Sat, na.rm = TRUE), o2_range = range(O2.Sat, na.rm = TRUE),

            `Hematocrit (%)` = median(Hematocrit, na.rm = TRUE), hem_range = range(Hematocrit, na.rm = TRUE),
            `Lactate (mmol/L)` = median(Lactate, na.rm = TRUE), lactate_range = range(Lactate, na.rm = TRUE),
            `Arginine (μmol/L)` = median(Arginine.umol.L, na.rm = TRUE), arg_range = range(Arginine.umol.L, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, round, 1) %>%

  mutate(`Age (years)` = paste0(`Age (years)`, " (", age_range[1], ", ", age_range[2], ")"),

         `Temperature (°C)` = paste0(`Temperature (°C)`, " (", temp_range[1], ", ", temp_range[2], ")"),
         `Heart rate` = paste0(`Heart rate`, " (", hr_range[1], ", ", hr_range[2], ")"),
         `Systolic BP` = paste0(`Systolic BP`, " (", sys_range[1], ", ", sys_range[2], ")"),
         `Diastolic BP` = paste0(`Diastolic BP`, " (", dia_range[1], ", ", dia_range[2], ")"),
         `Respiration rate` = paste0(`Respiration rate`, " (", rr_range[1], ", ", rr_range[2], ")"),
         `O~2~ saturation` = paste0(`O~2~ saturation`, " (", o2_range[1], ", ", o2_range[2], ")"),

         `Hematocrit (%)` = paste0(`Hematocrit (%)`, " (", hem_range[1], ", ", hem_range[2], ")"),
         `Lactate (mmol/L)` = paste0(`Lactate (mmol/L)`, " (", lactate_range[1], ", ", lactate_range[2], ")"),
         `Arginine (μmol/L)` = paste0(`Arginine (μmol/L)`, " (", arg_range[1], ", ", arg_range[2], ")")) %>% 
  dplyr::select(-c(age_range, temp_range, hr_range, sys_range, dia_range, rr_range, o2_range, hem_range, lactate_range, arg_range)) %>%

@@

```
**Table 1.** A summary of subject information and hematological data. Group patient data presented as median (range). Note: Sex not recorded for 3 CM patients; Lactate data removed for HV due to inconsistent data.

@@

```{r fig1, fig.height = 12}

@@

df_hbox_voi_long %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~variable, scales = "free_y", nrow = 4, ncol = 2) +

@@

**Figure 1.** Differences in variables measured by Oxiplex and associated α values among patient groups. THC levels were different between all three groups (Kruskal-Wallis adjusted p-value = 3.25x10^-6^; Dunn Test post-hoc adjusted p-values: CM vs HV = 1.66x10^-6^, CM vs UM = .032, HV vs UM = .033). Levels of cerebral O~2~ saturation did not differ among groups (Kruskal-Wallis adjusted p-value = .07). Note: removed one outlier from Deoxy Hb α (3.52).

### Figure 2: Logistic regression on CM vs UM

@@

```{r fig2a}
# remove healthy volunteers, Oxy Hb, and Deoxy Hb from analysis
df_hbox_log_voi <- df_hbox_voi %>% 
  dplyr::filter(Status != "HV") %>% 
  dplyr::select(-(2:5))

@@

```{r fig2b_prep, eval = FALSE}

@@

```{r fig2b_plot}

@@

**Figure 2** Logistic regression model discriminating CM vs UM patients using only THCα as the predictor. **(A)** PCA of patients with CM and UM as determined by NIRS variables. Color indicates patient status, ellipses show median and 95% confidence interval on the cluster. **(B)** Confusion matrix representing the results of the logistic regression model averaged across 1000 bootstrap resamples. Percent occurrence indicates percentage times that the event (shown by the x- and y-axes) occurred.

>>>>

# other numeric variables

df_clinical_res <- df_hbox %>%

@@

  group_by(Status) %>% 
  summarise(`Age (years)` = median(Age, na.rm = TRUE), age_q1 = quantile(Age, 0.25, na.rm = TRUE), age_q3 = quantile(Age, 0.75, na.rm = TRUE),

            `Temperature (°C)` = median(Temperature, na.rm = TRUE), temp_q1 = quantile(Temperature, 0.25, na.rm = TRUE), temp_q3 = quantile(Temperature, 0.75, na.rm = TRUE),
            `Heart rate` = median(HR, na.rm = TRUE), hr_q1 = quantile(HR, 0.25, na.rm = TRUE), hr_q3 = quantile(HR, 0.75, na.rm = TRUE),
            `Systolic BP` = median(BP.Systolic, na.rm = TRUE), sys_q1 = quantile(BP.Systolic, 0.25, na.rm = TRUE), sys_q3 = quantile(BP.Systolic, 0.75, na.rm = TRUE),
            `Diastolic BP` = median(BP.Diastolic, na.rm = TRUE), dia_q1 = quantile(BP.Diastolic, 0.25, na.rm = TRUE), dia_q3 = quantile(BP.Diastolic, 0.75, na.rm = TRUE),
            `Respiration rate` = median(RR, na.rm = TRUE), rr_q1 = quantile(RR, 0.25, na.rm = TRUE), rr_q3 = quantile(RR, 0.75, na.rm = TRUE),
            `O~2~ saturation` = median(O2.Sat, na.rm = TRUE), o2_q1 = quantile(O2.Sat, 0.25, na.rm = TRUE), o2_q3 = quantile(O2.Sat, 0.75, na.rm = TRUE),

            `Hematocrit (%)` = median(Hematocrit, na.rm = TRUE), hem_q1 = quantile(Hematocrit, 0.25, na.rm = TRUE), hem_q3 = quantile(Hematocrit, 0.75, na.rm = TRUE),
            `Lactate (mmol/L)` = median(Lactate, na.rm = TRUE), lac_q1 = quantile(Lactate, 0.25, na.rm = TRUE), lac_q3 = quantile(Lactate, 0.75, na.rm = TRUE),
            `Arginine (μmol/L)` = median(Arginine.umol.L, na.rm = TRUE), arg_q1 = quantile(Arginine.umol.L, 0.25, na.rm = TRUE), arg_q3 = quantile(Arginine.umol.L, 0.75, na.rm = TRUE)) %>% 
  mutate_if(is.numeric, round, 1) %>%

  mutate(`Age (years)` = paste0(`Age (years)`, " (", age_q1, ", ", age_q3, ")"),

         `Temperature (°C)` = paste0(`Temperature (°C)`, " (", temp_q1, ", ", temp_q3, ")"),
         `Heart rate` = paste0(`Heart rate`, " (", hr_q1, ", ", hr_q3, ")"),
         `Systolic BP` = paste0(`Systolic BP`, " (", sys_q1, ", ", sys_q3, ")"),
         `Diastolic BP` = paste0(`Diastolic BP`, " (", dia_q1, ", ", dia_q3, ")"),
         `Respiration rate` = paste0(`Respiration rate`, " (", rr_q1, ", ", rr_q3, ")"),
         `O~2~ saturation` = paste0(`O~2~ saturation`, " (", o2_q1, ", ", o2_q3, ")"),

         `Hematocrit (%)` = paste0(`Hematocrit (%)`, " (", hem_q1, ", ", hem_q3, ")"),
         `Lactate (mmol/L)` = paste0(`Lactate (mmol/L)`, " (", lac_q1, ", ", lac_q3, ")"),
         `Arginine (μmol/L)` = paste0(`Arginine (μmol/L)`, " (", arg_q1, ", ", arg_q3, ")")) %>% 
  dplyr::select(-c(age_q1, age_q3, 
                   temp_q1, temp_q3, hr_q1, hr_q3, sys_q1, sys_q3, dia_q1, dia_q3, rr_q1, rr_q3, o2_q1, o2_q3, 
                   hem_q1, hem_q3, lac_q1, lac_q3, arg_q1, arg_q3)) %>%

@@

```
**Table 1.** A summary of subject information and hematological data. Group patient data presented as median (quartile 1, quartile 3). Note: Sex not recorded for 3 CM patients; Lactate data removed for HV due to inconsistent data.

@@

```{r fig1}

@@

df_hbox_voi_long %>% dplyr::filter(variable %in% c("Oxy Hb", "Deoxy Hb", "Cerebral O2 sat", "THC")) %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~variable, scales = "free_y", nrow = 2, ncol = 2) +

@@

**Figure 1.** Shown are the variables measured by Oxiplex, plotted by patient status THC levels were different between all three groups (Kruskal-Wallis adjusted p-value = 3.25x10^-6^; Dunn Test post-hoc adjusted p-values: CM vs HV = 1.66x10^-6^, CM vs UM = .032, HV vs UM = .033). Levels of cerebral O~2~ saturation did not differ among groups (Kruskal-Wallis adjusted p-value = .07).

### Figure 2: Visual summary of α values associated with NIRS variables

```{r fig2}

df_hbox_voi_long %>% dplyr::filter(variable %in% c("Oxy Hb α", "Deoxy Hb α", "Cerebral O2 sat α", "THCα")) %>%

  ggplot(aes(x = Status, y = value)) +
  geom_violin(aes(color = Status)) +
  geom_jitter(position = position_jitter(0.2), shape = 1) +
  stat_summary(fun = "median", geom = "crossbar", aes(color = Status), size = 0.2, width = 0.5) +
  facet_wrap(~variable, scales = "free_y", nrow = 2, ncol = 2) +

**Figure 2** Shown are the α values associated with variables measured by the Oxiplex, plotted by patient status. Note: One outlier was removed from Deoxy Hb α (3.52).

### Figure 3: Logistic regression on CM vs UM

@@

```{r fig3a}
# remove healthy volunteers, Oxy Hb, and Deoxy Hb from analysis
df_hbox_log_voi <- df_hbox_voi %>% 
  dplyr::filter(Status != "HV") %>% 
  dplyr::select(-(2:5))

@@

```{r fig3b_prep, eval = FALSE}

@@

```{r fig3b_plot}

@@

**Figure 3** Logistic regression model discriminating CM vs UM patients using only THCα as the predictor. **(A)** PCA of patients with CM and UM as determined by NIRS variables. Color indicates patient status, ellipses show median and 95% confidence interval on the cluster. **(B)** Confusion matrix representing the results of the logistic regression model averaged across 1000 bootstrap resamples. Percent occurrence indicates percentage times that the event (shown by the x- and y-axes) occurred.
